---
title: "Lab1 - R basics"
author: "Klimov, Titova"
date: 'null'
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightLines: yes
      highlightStyle: github
  ioslides_presentation:
    df_print: paged
    widescreen: yes
  revealjs::revealjs_presentation:
    widescreen: yes
  beamer_presentation:
    latex_engine: xelatex
---

Part 1. RStudio as the IDE for R
========================================================

- Free software

- Object-oriented

- Available for all major operating systems (Mac OS, Linux, Windows)

- Can easily be extended through the use of user-defined functions

- Widely used in political science, statistics, econometrics, actuarial sciences, sociology, finance, etc.

For more details on authoring R presentations please visit <https://support.rstudio.com/hc/en-us/articles/200486468>.

---

Howto run code
========================================================

```{r}
summary(cars)
```

---

Slide With Plot
--------------------------------------------------------

```{r, echo=FALSE, fig.show='hold', fig.align='center'}
plot(cars)
```

---

Part 2. Intro  in R Language - python-alike console 
--------------------------------------------------------

```{r}
# A comment
5 # print a number
2+2 # calculations
x <- 10 # store
x # print
(x <- 11) # store and print
```

---

Vector - basic array
--------------------------------------------------------

**c**, which is short for concatenate (paste together) [https://cran.r-project.org/doc/contrib/Torfs+Brauer-Short-R-Intro.pdf]

```{r}
Ages <- c(32, 45, 15, 22, 29, 54, 35)  # store a specified vector to variable called "Ages"
Ages
```

---

Vector - basic array (2)
--------------------------------------------------------
Generate vector with **seq** and set it to **Participant_id**

```{r}
(Participant_id <- seq(1,7)) # define vector as a sequence (and print)
(Participant_id <- (1:7)) # makes the same
```

Seq could have additional parameters: seq(1,20, **by=2**)

```{r, echo=FALSE}
(seq(1,20, by=2)) # seq can have additional arguments
```

---

Operations with vectors
--------------------------------------------------------
Builtin statistics (mean, sd, statistical tests ...)

```{r}
mean(Ages)
length(Ages)
```

```{r, echo=FALSE}
TestingSkills <- c(5, 3, 4, 2, 5, 1, 4)
```

---

Matrices
--------------------------------------------------------
```{r}
mymatrix <- cbind(Participant_id, Ages, TestingSkills)
is.matrix(mymatrix); is.vector(TestingSkills) # to stack expressions in one line, use ";"
```

Let's check how much rows and colums we have
```{r}
nrow(mymatrix)
ncol(mymatrix)
```

---

Matrices is good but dataframe is better
--------------------------------------------------------
Didn't work in matrics:
```{r, error=TRUE}
mymatrix$Participant_id
```

But works in Dataframes:
```{r}
as.data.frame(mymatrix)$Participant_id
```

```{r}
mydataframe <- as.data.frame(mymatrix)
```

---

Building a plot
--------------------------------------------------------

```{r, echo=FALSE}
plot(mydataframe$Ages ,xlab="Ages"
     , mydataframe$TestingSkills, ylab="TestingSkills"
     , main="Proficiency"
     )
```

***

```{r eval=FALSE}
plot(mydataframe$Ages, mydataframe$TestingSkills
     , xlab="Ages", ylab="TestingSkills"
     , main="Proficiency"
     )
```

---

Lists, dataframes
--------------------------------------------------------

Can be viewed with:

* data
* str(data)
* View(data)
* DT(data)
* kable(data)

---

# Motor Trend Car Road dataset

**Description** The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models).

```{r, echo=FALSE}
#knitr::kable(mtcars, format = "html")
DT::datatable(
  mtcars,
  fillContainer = FALSE, options = list(pageLength = 5)
)
```

---

Task 1
----------------------------

**Load *mtcars_data* in a variable and display its**
- summary
- last 5 rows

---

Task 2
----------------------------

- Display the number of variables
- Display the number of records
- Add a new column "#" and number all the rows

---

Task 3
----------------------------

- Describe (Display all the info for) car #7
- Display the summary for column "qsec"
- Display first 6 values of "qsec"

---

Task 4
----------------------------
For first column display:
- values from 10 to 20 rows
- values for rows #4, #7, #9 in one line

---

Task 5
----------------------------

- Copy *kilometers per liter* to a new column "kml2"
- Declare new "kml2" variable from values of "miles per gallon", knowing that **_1 mile per liter = 0,425144 kilometer per liter_**

---

Task 6
----------------------------

- Display *True/False table* for all the rows of column *cyl*, checking if they are the same as 6
( + Find the *Median* of cylinder numbers)

---

Task 7
----------------------------

- Display top 5 rows, where *disp* is lesser than 150 

---

Task 8
----------------------------

- Display *car names*, for which *disp* is greater than 200

---

Task 9
----------------------------

- Count the *Mean Value* of *wt*, rounded to the nearest 10th.

---

Task 10
----------------------------

- Count those car models, that have 8 cylinders and horsepower greater than 150.
- From the subset above name a car with minimal *qsec*.
